---
title: "Project memo"
output: pdf_document
date: "2023-01-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='hide'}
setwd("~/Desktop/Third year/PSTAT 131/project/project data")
train <- read.csv("train.csv"); train
test <- read.csv("test.csv"); test
```

**Overview of dataset**

* I will be using an open source dataset found on Kaggle, and traced back to a Analytics Vidhya contest. 
  + See Kaggle data source [here](https://www.kaggle.com/datasets/altruistdelhite04/loan-prediction-problem-dataset?select=train_u6lujuX_CVtuZ9i.csv). 
  + See original problem [here](https://datahack.analyticsvidhya.com/contest/practice-problem-loan-prediction-iii/#ProblemStatement)
* The data file is split into a training dataset and a test dataset. 
  + The training dataset contains 614 observations and 13 variables.
  + The test datset contains 367 observations and 12 variables (predictors only).
* I will be working with a mix of quantitative and categorical variables. 
* There is some missing data. I plan to identify and remove the observations for which there is a missing record. 

**Research questions** 

* For my project, I want to build a machine learning model to predict whether a customer will be approved for a loan based on information in their application profile. 
* The predictors include Gender, Marital Status, Education, Number of Dependents, Income, Loan Amount, Credit History and others. 
* The response variable will be Loan_Status (Y/N), and represents loan eligibility.  
* I'm also interested in descriptive questions such as:
  + How many applicants are married based on historical data? How many have dependents? 
  + How do the incomes of applicants who live in different areas compare?
  + How do the incomes of employed applicants compare to those who are self-employed?
  + Is there a correlation between an applicant’s income and the loan amount they apply for? 
  + Is there bias? Does eduation level improve loan eligibility? Gender? 
  + How do the credit histories of male and female applicants compare? 
* Because my response is qualitative (classification) and I have no response for the test set (unsupervised), I will need to use a more complex method such as K-modes clustering. 
* However, I want to split my training set into a training set and a test set instead of using the one given in the data file. This way I can use supervised methods such as decision tree, naive bayes, or logistic regression. 
* For this project, I will focus on prediction and model selection. But if a functional form for f is avaliable (such as in logistic), some inference would be helpful to my analysis. 


**Project timeline** 

* I have already found my dataset. I plan to tidy my data and perform EDA on weeks 2 and 3. For the remainder of the quarter, I will focus on building my model and editing my final draft. 
* I expect to be done with my model around week 5-6 and my write up by week 7. 


